---
title: "MY first markdown file"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(ggplot2)
```

## Reading the Data
This dataset was a collection of favorite restaurants in the 10:30 am section of our Statistical Modeling class. The dataset includes name of the restaurant, price of average meal, last visit, cuisine, annual visit, and location.

```{r}
class_data <- read_csv("C:\\Users\\jp7nb\\Desktop\\fav_restaurants.csv")
```

## Plotting the Data

```{r}
ggplot(class_data, aes(cost_per_person)) + geom_bar()
```

This bar graph represents the average cost per person. This also shows the frequency of the cost.

```{r}
ggplot(class_data, aes(cost_per_person, yearly_visits)) + geom_line()
```

This line chart shows the yearly visits at the price point. This shows that as the food gets more expensive, the less frequent the yearly visits are. 

```{r}
ggplot(class_data, aes(yearly_visits)) + geom_bar()
```

This bar graph shows the frequency of visits at each restaurant. This shows that most students visit the restaurant around 0-10 times, with a large drop off over 10 visits. 





